Stable Hot Spot Analysis (Draft)

### 2017-06-16 paper-main.tex
text fixes

But, similar to a cluster analysis, there does exist a high dependency of the identified hotspots on the detection method and in particular the parametrization of this method. The identified areas as well as their shape can vary highly. This volatility can lead to a decrease in trust in the result or in suboptimal allocations of scarce resources. Therefore it is necessary to measure and evaluate the stability of a hotspot analysis as well as the different parametrizations. 
In our initial work~\cite{SoH-GI-Forum}, we introduced a method to measure the stability of hotspots, the \emph{stability of hotspots} metric (SoH) and showed its use on the basis of temperature data .   
Here, we build upon that work and examine in more detail the impact of the 
different instantiations of the most typical parameter. We use the well known 
Getis-Ord statistic \cite{Ord.1995}, the standard $G^*$, and a modification of 
this statistic, the focal $G^*$~\cite{SoH-GI-Forum}. Those parameter are the 
size of the study area (focal matrix), the detail of the resolution 
(aggregation level) and the size of the neighbourhood (weight matrix). By 
varying over these parameter, we can compare the stability for all possible 
combinations and isolate the effect of single parameter by aggregation over the 
other parameter. 
We evaluate on the well-known taxi data set to show the applicability on real world use cases as well as to enable a simple replication.

@@

Existing methods for determining hot spots are dependent on the parametrization
of the weight matrix as well as on the size of the study area. Intuitively,
increasing the size of a weight matrix has a "blurring" effect on the raster
(Fig.~\ref{def:TODO}a) whereas decreasing the size can be seen as a form of
"sharpening" (Fig.~\ref{def:TODO}b).

@@

%
%Consider the real-world example depicted in Fig.~\ref{fig:TempMaps}.
%%TODO: thermal flight -> NY taxi
%The temperature map of a morning thermal flight dataset 
%(Fig.~\ref{fig:TempMaps:a}) has been processed using $G^*$.
%
%%TODO: thermal flight -> NY taxi
%%TODO: need two images - G* and FocalG*
%\begin{figure}[htp]
%  \begin{subfigure}{1\linewidth}
%    \caption{Morning temperatures}
%    \includegraphics[width=\linewidth]{images/hotspot-rawtemp-1}
%    \label{fig:TempMaps:a}
%  \end{subfigure}
%  
%  \caption{
%    Karlsruhe city center.
%    Selected area of $2.4{\times}1.4\,km$.
%    Pixel size $5{\times}5\,m$.
%  }
%  \label{fig:TempMaps}
%\end{figure}

@@

That is

To be stable, one assumes that every parent has at
least one child and that each child has one parent. For a perfectly stable
interaction, it can be easily seen that the connection between parent and child
is a injective function and between child and parent a surjective function. To
measure the closeness of connection, we propose a metric called the
\emph{Stability of Hot spot} (SoH). It measures the deviation from a perfectly
stable transformation of resolutions.

@@

\section{Focal Getis-Ord} \label{sec:FocalGetisOrd}

Section is analog to \cite{SoH-GI-Forum} and incorporates the most important definitions for the computation.
We use the notation \FOCAL{R}{M}{op} to denote a focal
operation $op$ applied on a raster $R$ with a focal window determined by a
matrix $M$. This is rougly eqivalent to a command \verb|focal(x=R, w=M, fun=op)|
from package \emph{raster} in the R programming language \cite{cran:raster}.

\begin{definition}[$G^*$ function on rasters]
  \label{def:GenericGetisOrdFunc}

  The function $G^*$ can be expressed as a raster operation:

  \begin{displaymath}
    G^*(R, W, st) =
    \frac{
      \FOCAL{R}{W}{sum} - M*\sum_{w \in W}{w}
    }{
      S \sqrt{
        \frac{
          N*\sum_{w \in W}{w^2} - (\sum_{w \in W}{w})^2
        }{
          N - 1
        }
      }
    }
  \end{displaymath}
  \noindent where:
  \begin{itemize}

    \item $R$ is the input raster.

    \item $W$ is a weight matrix of values between 0 and 1.

    \item $st = (N, M, S)$ is a parametrization specific to a particular version
    of the $G^*$ function. (Def.\vref{def:StdGetisOrdVersion} and
    \vref{def:FocalGetisOrdVersion}).

  \end{itemize}
\end{definition}

\begin{definition}[Standard $G^*$ parametrization]
  \label{def:StdGetisOrdVersion}

  Computes the parametrization $st$ as global statistics for all pixels in the
  raster $R$:

  \begin{itemize}
    \item $N$ represents the number of all pixels in $R$.
    \item $M$ represents the global mean of $R$.
    \item $S$ represents the global standard deviation of all pixels in $R$.
  \end{itemize}
\end{definition}

\begin{definition}[Focal $G^*$ parametrization]
  \label{def:FocalGetisOrdVersion}

  Let $F$ be a boolean matrix such that: $all(dim(F) \geq dim(W))$.
  This version uses focal operations to compute per-pixel statistics given by 
  the focal neighbourhood $F$ as follows:

  \begin{itemize}

    \item $N$ is a raster computed as a focal operation \FOCAL{R}{F}{sum}. Each
    pixel represents the number of pixels from $R$ convoluted with the matrix
    $F$.

    \item $M$ is a raster computed as a focal mean \FOCAL{R}{F}{mean}, thus each
    pixel represents a mean value of its $F$-neighbourhood.

    \item $S$ is a raster computed as a focal standard deviation
    \FOCAL{R}{F}{sd}, thus each pixel represents a standard deviation of its
    $F$-neighbourhood.

  \end{itemize}

\end{definition}

@@

\begin{definition}[Evaluation Run]
  \label{def:EvalRun}
  We define a single evaluation run as a tuple:
  \begin{displaymath}
    E = (V, m, p, w)
  \end{displaymath}
  \noindent where:
  \begin{itemize}

    \item $V$ is the input dataset of points, representing the taxi dropoffs in 
    our case.

    \item $m$ is the metric used, in our case either \SOHUP or \SOHDOWN.

    \item $p$ represents the pixel size for aggregating points from $V$, 
    e.g. $100 \times 100$ meters.

    \item $w$ represents the size of a weight matrix.
    In our case, we chose a weight matrix depicted in 
    Figure~\ref{fig:ExampleMatrices}(a) for both the G* and FocalG* cases.
  \end{itemize}
\end{definition}

%Fig.~\ref{fig:ComparisonMorning} and Fig.~\ref{fig:ComparisonEvening} show
%Standard and Focal $G^*$ computations for both morning and eventing datasets
%with weight matrix $W$ of size 3, 5, 7, 9, 15 and 31. In this figures, when 
%computing Focal
%$G^*$, the focal matrix $F$ has a constant size of $61{\times}61$ cells. 
%Example weight matrix $W$ and focal matrix $F$ are depicted in 
%Fig.~\ref{fig:ExampleMatrices}.

@@

Therefore we would calculate $10 \cdot 10 \cdot 5 + 10 \cdot 5=550$ results for each dimension.
The following conditions must hold dim(F)>dim(W) and dim(F)<dim(R) which leads e.g. to 460 results in the z dimension.
We vary over the weight matrix W, the focal matrix F and the aggregation level, as motivated in the introduction.
To compare the impact each of these parameters has we compute all variations over two parameter and holding one parameter fix. We then calculate the mean as well as the standard deviation for the SoH for the fixed value based on the two other parameters. This allows us to isolate the impact of the variation in the single parameter. The results are then plotted in Fig. 1-6, one graphic for each fixed parameter and the direction of the SoH.
The hotspots to use the SoH on are computed by the focal $G^*$ and standard $G^*$ method. The focal $G^*$ allows us, to isolate the impact that the size of the study area has on the stability of the results without recomputing the total dataset.

@@

\subsection{Clumping}
The SoH needs a parent and a child. Clusters can be calculated different. For $G^*$ a sigmoid can be used. Values above 2.58 and under -2.58 means in less then 
1\% of the cases you are wrong. The values of Focal $G^*$ depend on the focal size F. Therefore, we decide all values in the top quantile belong to a clusters.
For clustering we grouped the neighbourhood with values in the top quantile to one cluster. 
Neigbours are are pixels with touching edges which leads for every pixel to 8 neigbours.
The clustering can be expressed in R with clumping, TODO which package and parametrisations.
In Fig.~\ref{fig:Clumping} we visualized the clusters. The clusters are marked in different colours yellow (child) and violet (parent).
It can seen that the results from $G^*$ and Focal $G^*$ differ. Focal $G^*$ has more smaller clusters which are distributed over the complete raster.
Not overlapping clusters should be not stable because if we changed the parametrisation we want that the clusters do not disappear or appear. 
The SoH defines in more detail, which changes for the clusters are stable and which changes indicate that the clusters are instable.

\begin{figure*}[htp]
  \centering

  \begin{subfigure}{.35\linewidth}
    \caption{G* clustering difference}
    \includegraphics[width=\linewidth]{images/gen-metric-example-1}
    \label{fig:ClustDiffGstar}
  \end{subfigure}
  \hspace{3em}
  \begin{subfigure}{.35\linewidth}
    \caption{Focal G* clustering difference}
    \includegraphics[width=\linewidth]{images/gen-metric-example-2}
    \label{fig:ClustDiffFocalGstar}
  \end{subfigure}

  \caption{
    Example clustering for G* and Focal G* used as a input for SoH computation.
  }
  \label{fig:Clumping}
\end{figure*}

@@

The results can be found in figure ~\ref{fig:Zoom}. It can be seen that with an
increase of the aggregation level the values increase. TODO if clusters
increase/decrease For the upward property, $G^*$ is always better than Focal
$G^*$, compared to the results from figure \ref{fig:SoHZoom}. Focal $G^*$
reaches the same result as $G^*$ at aggregation level~5 . For the downward
property, Focal $G^*$ leads to better results at aggregation level~5.

@@

\subsection{Blur}
We also made an example for the variation of the weight size W.
Blur is a slice plane through our three dimensional space. 
W is changed from 7 to 29 with a stepsize of 2, F and the aggregation level is fixed.
For blur we also included the negative clusters which resulted in a surprising result.
\begin{definition} Blur run:
\begin{displaymath}
\SOHUP( G^*(R_z, W_{7+i}, st, F), G^*(R_{z}, W_{7+i+2}, st, F) )
\end{displaymath}
\begin{displaymath}
\SOHDOWN( G^*(R_z, W_{7+i}, st, F), G^*(R_{z}, W_{7+i+2}, st, F) )
\end{displaymath}
\begin{displaymath}
z =3, i \in [1,6],dim(F)=41, G^* \in [Standard, Focal]
\end{displaymath}
\end{definition}
The blur results are plotted in Fig.~\ref{fig:Blur}. The weight matrix is shown in the first row, with increasing size.
The scond and thrid row sho how $G^*$ and Focal $G^*$ changes based on W.
This example is a slice plane of the weight dimension ploted in figure ~\ref{fig:SoHWeight}.
It can be seen that increase of the weight size leads in general to better results for Focal $G^*$. 
$G^*$ compared to the mean of our whiskerplot does not fit in the picture because it leads to better results than $G^*$.

>>>>

But, similar to a cluster analysis, there does exist a high dependency of the identified hotspots on the detection method and in particular the parametrization of this method. The identified areas as well as their shape can vary highly. This volatility can lead to a decrease in trust in the result or in suboptimal allocations of scarce resources. Therefore it is necessary to measure and evaluate the stability of a hotspot analysis as well as the different parametrizations. 
In our initial work~\cite{SoH-GI-Forum}, we introduced a method to measure the stability of hotspots, the \emph{stability of hotspots} metric (SoH) and showed its use on the basis of temperature data .   
Here, we build upon that work and examine in more detail the impact of the 
different instantiations of the most typical parameter. We use the well known 
Getis-Ord statistic \cite{Ord.1995}, the standard $G^*$, and a modification of 
this statistic, the Focal $G^*$~\cite{SoH-GI-Forum}. Those parameter are the 
size of the study area (focal matrix), the detail of the resolution 
(aggregation level) and the size of the neighbourhood (weight matrix). By 
varying over these parameter, we can compare the stability for all possible 
combinations and isolate the effect of single parameter by aggregation over the 
other parameter. 
We evaluate on the well-known taxi data set to show the applicability on real world use cases as well as to enable a simple replication.

@@

Existing methods for determining hot spots are dependent on the parametrization
of the weight matrix as well as on the size of the study area. Intuitively,
increasing the size of a weight matrix has a "blurring" effect on the raster
(Figure~\ref{fig:BlurExample:a}) whereas decreasing the size can be 
seen as a form of "sharpening" (Figure~\ref{fig:BlurExample:b}).

@@

%
%Consider the real-world example depicted in Figure~\ref{fig:TempMaps}.
%%TODO: thermal flight -> NY taxi
%The temperature map of a morning thermal flight dataset 
%(Figure~\ref{fig:TempMaps:a}) has been processed using $G^*$.
%
%%TODO: thermal flight -> NY taxi
%%TODO: need two images - G* and FocalG*
\begin{figure}[htp]
  \centering
  \begin{subfigure}{.3\linewidth}
    \caption{Small matrix}
    \includegraphics[width=\linewidth]{images/gen-raw-blur-gstar-2}
    \label{fig:BlurExample:a}
  \end{subfigure}
  \begin{subfigure}{.3\linewidth}
    \caption{Large matrix}
    \includegraphics[width=\linewidth]{images/gen-raw-blur-gstar-8}
    \label{fig:BlurExample:b}
  \end{subfigure}
  \caption{
    Example of G* statictics computed for the same raster with different weight 
    matrix sizes.
  }
\end{figure}

@@

Stability assumes that every parent has at least one child and that each child
has one parent. For a perfectly stable interaction, it can be easily seen that
the connection between parent and child is a injective function and between
child and parent a surjective function. To measure the stability,
we propose a metric called the \emph{Stability of Hot spot} (SoH). It measures
the deviation from a perfectly stable transformation between the parent and 
chaild rasters (like frames of an animation).

@@

%\section{Focal Getis-Ord}
%\label{sec:FocalGetisOrd}
%
%In paper~\cite{SoH-GI-Forum}, we have defined the measure Focal G*.
%We use the notation \FOCAL{R}{M}{op} to denote a focal
%operation $op$ applied on a raster $R$ with a focal window determined by a
%matrix $M$. This is rougly eqivalent to a command \verb|focal(x=R, w=M, 
%fun=op)|
%from package \emph{raster} in the R programming language \cite{cran:raster}.

%\begin{definition}[$G^*$ function on rasters]
%  \label{def:GenericGetisOrdFunc}
%  
%  The function $G^*$ can be expressed as a raster operation:
%  
%  \begin{displaymath}
%    G^*(R, W, st) =
%    \frac{
%      \FOCAL{R}{W}{sum} - M*\sum_{w \in W}{w}
%    }{
%      S \sqrt{
%        \frac{
%          N*\sum_{w \in W}{w^2} - (\sum_{w \in W}{w})^2
%        }{
%          N - 1
%        }
%      }
%    }
%  \end{displaymath}
%  \noindent where:
%  \begin{itemize}
%    
%    \item $R$ is the input raster.
%    
%    \item $W$ is a weight matrix of values between 0 and 1.
%    
%    \item $st = (N, M, S)$ is a parametrization specific to a particular 
%version
%    of the $G^*$ function. (Def.\vref{def:StdGetisOrdVersion} and
%    \vref{def:FocalGetisOrdVersion}).
%    
%  \end{itemize}
%\end{definition}
%
%\begin{definition}[Standard $G^*$ parametrization]
%  \label{def:StdGetisOrdVersion}
%  
%  Computes the parametrization $st$ as global statistics for all pixels in the
%  raster $R$:
%  
%  \begin{itemize}
%    \item $N$ represents the number of all pixels in $R$.
%    \item $M$ represents the global mean of $R$.
%    \item $S$ represents the global standard deviation of all pixels in $R$.
%  \end{itemize}
%\end{definition}
%
%\begin{definition}[Focal $G^*$ parametrization]
%  \label{def:FocalGetisOrdVersion}
%  
%  Let $F$ be a boolean matrix such that: $all(dim(F) \geq dim(W))$.
%  This version uses focal operations to compute per-pixel statistics given by 
%  the focal neighbourhood $F$ as follows:
%  
%  \begin{itemize}
%    
%    \item $N$ is a raster computed as a focal operation \FOCAL{R}{F}{sum}. Each
%    pixel represents the number of pixels from $R$ convoluted with the matrix
%    $F$. 
%    
%    \item $M$ is a raster computed as a focal mean \FOCAL{R}{F}{mean}, thus 
%each
%    pixel represents a mean value of its $F$-neighbourhood.
%    
%    \item $S$ is a raster computed as a focal standard deviation
%    \FOCAL{R}{F}{sd}, thus each pixel represents a standard deviation of its
%    $F$-neighbourhood.
%    
%  \end{itemize}
%  
%\end{definition}

@@

%\begin{definition}[Evaluation Run]
%  \label{def:EvalRun}
%  We define a single evaluation run as a tuple:
%  \begin{displaymath}
%    E = (V, m, p, w)
%  \end{displaymath}
%  \noindent where:
%  \begin{itemize}
%  
%    \item $V$ is the input dataset of points, representing the taxi dropoffs 
%in 
%    our case.
%    
%    \item $m$ is the metric used, in our case either \SOHUP or \SOHDOWN.
%    
%    \item $p$ represents the pixel size for aggregating points from $V$, 
%    e.g. $100 \times 100$ meters.
%    
%    \item $w$ represents the size of a weight matrix.
%    In our case, we chose a weight matrix depicted in 
%    Figure~\ref{fig:ExampleMatrices}(a) for both the G* and FocalG* cases.
%  \end{itemize}
%\end{definition}

%Figure~\ref{fig:ComparisonMorning} and Figure~\ref{fig:ComparisonEvening} show
%Standard and Focal $G^*$ computations for both morning and eventing datasets
%with weight matrix $W$ of size 3, 5, 7, 9, 15 and 31. In this figures, when 
%computing Focal
%$G^*$, the focal matrix $F$ has a constant size of $61{\times}61$ cells. 
%Example weight matrix $W$ and focal matrix $F$ are depicted in 
%Figure~\ref{fig:ExampleMatrices}.

@@

Therefore we would calculate $10 \cdot 10 \cdot 5 + 10 \cdot 5=550$ results for each dimension.
The following conditions must hold $dim(F) > dim(W)$ and $dim(F) < dim(R)$ 
which leads e.g. to 460 results in the z-dimension.
We vary over the weight matrix W, the focal matrix F and the aggregation level, as motivated in the introduction.
To compare the impact each of these parameters has we compute all variations 
over two parameter and holding one parameter fix. We then calculate the mean as 
well as the standard deviation for the SoH for the fixed value based on the two 
other parameters. This allows us to isolate the impact of the variation in the 
single parameter. The results are then plotted in 
Figures~\ref{fig:SoHFocal}--\ref{fig:SoHZoom}, one image for each fixed 
parameter and the direction of the SoH.
The hotspots to use the SoH on are computed by the focal $G^*$ and standard $G^*$ method. The focal $G^*$ allows us, to isolate the impact that the size of the study area has on the stability of the results without recomputing the total dataset.

@@

\subsection{Clumping}

The SoH needs a parent and a child. Clusters can be calculated different. For
$G^*$ a sigmoid can be used. Values above 2.58 and under -2.58 means in less
then 1\% of the cases you are wrong. The values of Focal $G^*$ depend on the
focal size F. Therefore, we decide all values in the top quantile belong to a
clusters. For clustering we grouped the neighbourhood with values in the top
quantile to one cluster. Neigbours are are pixels with touching edges which
leads for every pixel to 8 neigbours. Fig.~\ref{fig:Clumping} depicts the 
clusters using different colours. In this figure, Focal~$G^*$ has smaller 
clusters which are distributed over the complete raster.

\begin{figure}[htp]
  \centering

  \begin{subfigure}{.7\linewidth}
    \caption{G* clustering difference}
    \includegraphics[width=\linewidth]{images/gen-metric-example-1}
    \label{fig:ClustDiffGstar}
  \end{subfigure}
  \hspace{3em}
  \begin{subfigure}{.7\linewidth}
    \caption{Focal G* clustering difference}
    \includegraphics[width=\linewidth]{images/gen-metric-example-2}
    \label{fig:ClustDiffFocalGstar}
  \end{subfigure}

  \caption{
    Example clustering for G* and Focal G* used as a input for SoH computation.
  }
  \label{fig:Clumping}
\end{figure}

@@

The results can be found in Fig.~\ref{fig:Zoom}. It can be seen that with an
increase of the aggregation level the values increase. TODO if clusters
increase/decrease For the upward property, $G^*$ is always better than Focal
$G^*$, compared to the results from figure \ref{fig:SoHZoom}. Focal $G^*$
reaches the same result as $G^*$ at aggregation level~5 . For the downward
property, Focal $G^*$ leads to better results at aggregation level~5.

@@

\subsection{Blur}
We also made an example for the variation of the weight size W.
Blur is a slice plane through our three dimensional space. 
W is changed from 7 to 29 with a stepsize of 2, F and the aggregation level is fixed.
For blur we also included the negative clusters which resulted in a surprising result.
\begin{definition} Blur run:
\begin{displaymath}
\SOHUP( G^*(R_z, W_{7+i}, st, F), G^*(R_{z}, W_{7+i+2}, st, F) )
\end{displaymath}
\begin{displaymath}
\SOHDOWN( G^*(R_z, W_{7+i}, st, F), G^*(R_{z}, W_{7+i+2}, st, F) )
\end{displaymath}
\begin{displaymath}
z =3, i \in [1,6],dim(F)=41, G^* \in [Standard, Focal]
\end{displaymath}
\end{definition}
The blur results are plotted in Figure~\ref{fig:Blur}. The weight matrix is
shown in the first row, with increasing size. The scond and thrid row sho how
$G^*$ and Focal $G^*$ changes based on W. This example is a slice plane of the
weight dimension ploted in figure ~\ref{fig:SoHWeight}. It can be seen that
increase of the weight size leads in general to better results for Focal $G^*$.
$G^*$ compared to the mean of our whiskerplot does not fit in the picture
because it leads to better results than $G^*$.
